Different rules for binocular combination of luminance in cortical and subcortical pathways

### 2022-09-26 manuscript.Rmd
Testing commit
Added introduction

---
title: "Different rules for binocular combination of luminance in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Alex R. Wade &  Daniel H. Baker (+Myat? +Joel?)"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  pdf_document:
    toc: no
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

@@

>>>>

---
title: "Different rules for binocular combination of luminance in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Alex R. Wade &  Daniel H. Baker (+Myat? +Joel?)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  word_document: default
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

BinocularBinocular combination provides a higher visual sensitivity than monocular viewing. This superiority is known as binocular summation and is defined by the binocular summation ratio (BSR), which was originally believed to be around $\sqrt{2}$ ($\approx1.4$) for grating stimuli at detection threshold [@Campbell1965]. In other words, a monocular stimulus can elicit the same response as a binocular stimulus if it has a contrast that is 1.4 times higher. This led Legge to develop a widely accepted explanation that used quadratic summation to describe binocular combination [@Legge1984]: monocular signals from the right (R) and the left (L) eyes are squared before being summed together and the binocular response (B) is given by the square root of the output (B = $\sqrt{R^2 + L^2}$, when R and L are equal to 1, the output is $\sqrt{2}$). However, subsequent research has shown that these explanations are not fully adequate to account for binocular summation. Both of these accounts constitute single channel models, and this type of model has been shown to not being able to account for contrast detection in the presence of noise [@Anderson1989]. Moreover, more recent research has shown that the summation ratio can vary greatly between $\sqrt{2}$ and 2 depending on factors such as the spatial and temporal frequency of a stimulus or the sensitivity difference between the eyes [@Baker]. These observations led to the development of multistage gain control models, which combine binocular summation and interocular suppression, and can account for contrast matching, detection and discrimination for spatial contrast [@Ding2006; @Meese2006].
In general, it seems that the mechanisms behind binocular combination have been thoroughly studied, as have been the anatomical pathways behind it: light enters the eye though the pupil and signals are sent from the left and right retinae to the primary visual cortex, remaining anatomically isolated while passing through the lateral geniculate nucleus (LGN) until they reach V1, where they are binocularly combined [@Purves2008]. However, there is an eye component that is often underestimated in its role to determine the quality of visual information: the pupil. The pupils are openings found in the centre of the eyes that appear to be black and allow light to enter the eyes. Their size determines how much light will reach the retina and it is usually determined by the ambient levels of light: in brightness the pupils will constrict and in darkness they will dilate. This is known as the pupillary light response (PLR). The anatomical pathways that regulate this response are well understood and are very clearly and extensively described in the literature [@Angée2021; @Mathôt2018; @McDougal2010; @Wang2015]. However, they are anatomically distinct from the LGN-V1 pathway meaning that binocular combination occurs separately in anatomically distinct pathways. Given this, not much is known about the computational processes behind the PLR except for some evidence of binocular interaction. The presence of a consensual response in one eye when the other is being stimulated, and the presence of convergence (one pupil responds to illumination in either retina) and divergence (both pupils respond to illumination of one retina) [@Wyatt1981] are evidence of this binocular interaction.
With this in mind we designed an experiment that simultaneously recorded electrophysiological and pupillometric responses to investigate the combination of flickering light signals in both the visual cortex and the pupils. The results should offer new insight about basic neural circuits and information on how they might be affected in clinical disorders of vision (e.g. amblyopia). Based on previous literature, we expected to find a non-linear combination of the responses in visual cortex, as described by the gain control mechanisms, and a more linear combination of the responses and a greater binocular response at the level of the pupils.
To follow up on the results that we obtained, we decided to perform a contrast matching experiment to investigate whether perception of flickering light is consistent with the results observed in the cortical pathways (visual cortex) or the subcortical pathways (pupils). Matching is a paradigm in which the perceived brightness of a standard stimulus is matched to that of a target stimulus. In the latter, the interocular ratios of luminance are varied to obtain an equibrightness curve. Previous literature has used this paradigm to investigate the binocular fusion of static stimuli and the temporal combination of spatial flickering of spatial increments (a bright target on a dark background) and decrements (a dark target on a bright background) [@Anstis1998; @Levelt1965]. For spatial increments, it was found that binocular fusion seems to follow approximately linear combination rules. This means that, for a monocular stimulus to elicit the same response as a binocular stimulus, the former needs to have twice the signal of the latter. On the other hand, spatial decrements follow a winner-takes-all pattern. This means that the observer is seeing what the eye that is receiving the strongest signal is seeing.
Our experiment used different stimuli than the ones used by Anstis and Ho: in the binocular fusion experiment, their stimuli were not flickering and, in the flicker experiment, they were always shown in both eyes. In our experiment, in some conditions, the flicker was shown to only one eye. Moreover, they were looking at the temporal fusion of the flicker while we focussed on the binocular fusion of the flicker. Based on this and on the results from our first experiment, we expected to find a near linear summation of the responses.

@@

Figure \@ref(fig:pupildata)a
